---
title: My dynamic report
output: html_fragment
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, error=FALSE)
```

```{css}
body {
  background-color: #f8f9fa;
}

h1 {
  color: #ff5470;
  font-family: Arial;
}

p {
  color: #333333;
  background-color: #ffe4e1;
  padding: 10px;
  font-weight: bold;
}

```

## Tracking Phone Use Patterns

This dynamic report summarises my observations of phone use over time. 
The data was collected through a simple form, focusing on how long the phone was used, the main activity, and whether the usage had a clear purpose. 

By analysing this data, I aim to identify patterns in phone usage behaviour and compare how different activities and purposes relate to the amount of time spent on the phone.

```{r}
library(tidyverse)

logged_data <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vRnJ1vMcHVjJFKUIL4sHH_HXELf-HATASGe4N2tPIsy2l1EPVRWWMJB1366lz7e0c38Sf43m5wSx2BA/pub?gid=1914332814&single=true&output=csv")

latest_data <- logged_data %>%
  rename(
    purpose = `Did you have a clear purpose for using your phone?`,
    activity = `What was your main activity?`,
    time = `Approximately how many minutes did you use your phone? Please round to the nearest 100 minutes.`
  )
#summary values
mean_time <- mean(latest_data$time, na.rm = TRUE)
max_time <- max(latest_data$time, na.rm = TRUE)

```


```{r}
#bar charts 1
ggplot(latest_data) +
  geom_bar(aes(x = time, fill = activity), position = "dodge") +
  labs(
    title = "Phone Usage Time by Activity",
    subtitle = "Comparing time spent across different activities",
    x = "Time (minutes)",
    y = "Number of users"
  )
```

```{r}
#bar charts 2
ggplot(latest_data) +
  geom_bar(aes(x = time, fill = purpose), position = "dodge") +
  labs(
    title = "Phone Usage Time by Purpose",
    subtitle = "Comparing time spent by whether phone use had a clear purpose",
    x = "Time (minutes)",
    y = "Number of users"
  )
```

The average phone use time was `r round(mean_time, 2)` minutes.

The maximum phone use time was `r max_time` minutes.

The chart 1 shows that most phone use is for entertainment, with higher usage times compared to other activities. 
Study-related use is less frequent and usually occurs at shorter durations.

The chart 2 shows that phone use without a clear purpose tends to occur over a wider range of time and can be longer. 
In contrast, purposeful use is more concentrated at moderate time levels.
